---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Dhruv Kapadia"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.

```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```

If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?
```{r}
colnames(covid_data_race)[2]="State"
total <- merge(covid_data_orig, covid_data_race, by="State")

total |> 
  filter(State == "NJ") |> 
ggplot() +
geom_line(aes(x = Date, y = (Deaths_Asian/Asian_population)*1000), color = "blue") +
geom_line(aes(x = Date, y = (Deaths_Black/Black_population)*1000), color = "purple") +
geom_line(aes(x = Date, y = (Deaths_White/White_population)*1000), color = "red") +
geom_line(aes(x = Date, y = (Deaths_Latinx/Latinx_population)*1000), color = "orange") + 
  labs(title = "Cumulative Death Rate Per Capita Amongst Racial Groups in New Jersey", y = "Cumulative Death Rate per 1,000 People") + 
  theme_bw()
```
The graph above captures the cumulative rates of COVID-19-related deaths by race per 1,000 people in 
New Jersey for the state's four major racial groups. As demonstrated in the graph above, Black New 
Jerseyans have the highest cumulative death rate per capita relative to any other racial group. 
Conversely, Asian New Jerseyans have the lowest cumulative death rate per capita relative to any other
racial group. It is also important to note that White, Latinx, and Black New Jerseyans have a 
cumulative death rate increasing at approximately the same rate since January. 

2. Are there any conclusions you can draw?
As demonstrated in the graph above, Black New Jerseyans have the worst observable cumulative death 
rate associated to COVID-19 relative to other races. However, we cannot conclude that the true 
cumulative death rate of Black New Jerseyans is significantly worse than other racial groups without 
hypothesis testing. We also cannot explain why Black New Jerseyans were worst affected by COVID using 
the graph above. However, we can infer a myriad of reasons, from a lack of public health 
infrastructure to low vaccination rates in the Black community (due to historical mistrust with the 
medical community), behind why the Black cumulative death rate is significantly worse. Likewise, we 
cannot conclude why Asian New Jerseyans have the best outcomes with COVID-19 related deaths relative 
to other racial groups.

3. What were you not able to do due to not having the R knowledge?
I was unable to more effectively merge multiple data sets (I ended up using the more inefficient merge
function) in order to compare my x and y variables. I also was unable to break down the data in a more
granular way, as I had initially hoped to isolate post-January months rather than all of the available
dates in the datasets. More generally, I was unable to make the graph aesthetically pleasing. While I
randomly assigned colors to each racial group, I would've like to experiment more with color, line
thickness, and line style.

4. What other data would be useful to better understand this data?
It would have been helpful to have death counts per day by race rather than a cumulative death count.
This would have given me the changes in death rate over time, which could have been interesting to
measure when considering the effect of vaccines within certain communities.
